Python as a machine learning language: experiences from the \sklearn project

### 2013-06-11 paper.tex
some general principles for the API

% working title, please change
\title{Python as a machine learning language:
       experiences from the scikit-learn project}

@@

\begin{abstract}
The \sklearn project is an increasingly popular open source machine
learning library designed to be simple and efficient, accessible to non-experts,
and reusable in various contexts. In this paper, we present and discuss our
design choices for the application programming interface (API) of the project.
In particular, we describe the simple and elegant interface shared by all
learning and processing units in the library and then discuss its advantages in
terms of composition and reusability. The paper also comments on implementation
details specific to the Python ecosystem and analyzes obstacles faced by users
and developers of the library.
\end{abstract}

@@

\section{The scikit-learn project}

@@

Started in 2009, \sklearn is developed by an international team
of a dozen of core developers, mostly researchers from various fields
(e.g., machine learning, natural language processing, or neuroscience).
The project also greatly benefits from many occasional
contributors proposing small code or documentation fixes or improvements.
Development proceeds on GitHub (\url{https://github.com/scikit-learn}),
a platform which greatly facilitates this kind of collaboration.
Because of the large number of developers,
particular emphasis is put on keeping the project maintainable.
In particular, code must follow specific
quality guidelines, such as style consistency and unit-test coverage. In
addition, complete documentation and examples are required for all
features.
Finally, additions to the code base must pass code review
by at least two developers not involved in the implementation
of the proposed addition.

@@

@@

% Construction and application of an estimator object
% proceeds in a way that is somewhat reminiscent of partial function application.
% The constructor, \texttt{\_\_init\_\_}\footnote{
%   Strictly speaking, \texttt{\_\_init\_\_} is Python's
%   class \textit{initialization} method,
%   while \texttt{\_\_new\_\_} is the constructor.
%   The distinction is not relevant to the present description of scikit-learn.}
% takes a set of named (hyper-)parameters
% and attaches these to the estimator as public attributes.\footnote{
%   The Python language does not actually distinguish between public and private
%   attributes and methods,
%   but by convention, any identifier prefixed with an underscore (\texttt{\_})
%   is considered ``a non-public part of the API\ldots
%   an implementation detail and subject to change without notice.''
%   \Citep{pythontut}
%   }
% It never performs actual learning, nor does it even see a training set.

>>>>

% working title, please change
\title{Python as a machine learning language:
       experiences from the \sklearn project}

@@

\begin{abstract}
The \sklearn project is an increasingly popular open source machine
learning library designed to be simple and efficient, accessible to non-experts,
and reusable in various contexts. In this paper, we present and discuss our
design choices for the application programming interface (API) of the project.
In particular, we describe the simple and limited interface shared by all
learning and processing units in the library and then discuss its advantages in
terms of composition and reusability. The paper also comments on implementation
details specific to the Python ecosystem and highlights obstacles faced by users
and developers of the library.
\end{abstract}

@@

\section{The \sklearn project}

@@

Started in 2009, \sklearn is developed by an international team
of several dozen volunteers, mostly researchers from various fields
(e.g., machine learning, natural language processing, or neuroscience).
The project also greatly benefits from many occasional
contributors proposing small code or documentation fixes or improvements.
Development proceeds on GitHub (\url{https://github.com/scikit-learn}),
a platform which greatly facilitates this kind of collaboration.
Because of the large number of developers,
particular emphasis is put on keeping the project maintainable.
In particular, code must follow specific
quality guidelines, such as style consistency and unit-test coverage. In
addition, complete documentation and examples are required for all
features.
Finally, additions to the code base must pass code review
by at least two developers not involved in the implementation
of the proposed addition.

@@

The API is designed to adhere to the following broad principles:
\begin{description}
  \item[Inspectability.]
       \sklearn objects store their constructor parameters
       as well as parameter values determined by learning algorithms
       as public attributes in a documented format.
       (The Python language does not actually distinguish
       between public and private attributes of objects,
       but by convention, any identifier
       prefixed with an underscore (\texttt{\_})
       is considered ``a non-public part of the API''; \citealp{pythontut}.)
  \item[No proliferation of classes.]
       Learning algorithms are objects of custom classes,
       but datasets and hyperparameter names and values,
       are represented as standard Python, NumPy and SciPy classes
       rather than custom ones whenever feasible.
       Custom classes adhere to Python conventions and protocols
       for constructs such as iteration over the ``folds''
       in a $k$-fold cross validation.
       This should keep \sklearn easy to use
       and easy to combine with other libraries.
       % XXX do we need to elaborate more on this, or do we expect everyone
       % to be an experienced Java/adapter pattern hater?
  \item[Composability.]
       Many machine learning tasks are expressible
       as \textit{pipelines} of transformations to data.
       Other tasks are naturally viewed
       as modifications to other tasks;
       e.g., multi-label classification
       can be reduced to multiple binary classification tasks.
       Accordingly, multi-label classification
       is implemented by a class that takes an existing classifier object
       as input and modifies its function.
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       a default value for that parameter must be defined by the library
       if at all possible,
       and the default should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}

@@

% Construction and application of an estimator object
% proceeds in a way that is somewhat reminiscent of partial function application.
% The constructor, \texttt{\_\_init\_\_}\footnote{
%   Strictly speaking, \texttt{\_\_init\_\_} is Python's
%   class \textit{initialization} method,
%   while \texttt{\_\_new\_\_} is the constructor.
%   The distinction is not relevant to the present description of scikit-learn.}
% takes a set of named (hyper-)parameters
% and attaches these to the estimator as public attributes.
% It never performs actual learning, nor does it even see a training set.
